---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: true
  html_document: default
geometry: margin=1.5in
---
```{r eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE}
fig.height = 2.5
fig.width = 1.5 * fig.height
font_size = 8

knitr::opts_chunk$set(dev = 'tikz',
			   fig.height = fig.height, fig.width = fig.width,
			   fig.align='center', fig.pos = 't',
			   sanitize = TRUE,
			   echo = FALSE,
			   warning = FALSE, message = FALSE
			   )
options(digits = 1, xtable.comment = FALSE)
```

# Study 1: Altmetrics #

The term "altmetrics" has two distinct meanings.  Small-a "altmetrics" is a portmanteau of "alternative bibliometrics," that is, alternatives to such metrics as paper citation counts, journal impact factors, and researcher H-indices.  These standard bibliometrics are designed to measure the impact of research within the scholarly community, and thus are generally not useful for assessing social impacts.  Consequently, there is a substantial amount of active work on developing altmetrics that are relevant to social impact.  *[cite Britt etc.]*

Large-a "Altmetrics" refers to "Altmetric.com," a UK-based company that develops an integrated set of small-a altmetrics and makes them available programmatically using a web-based API [application programming interface].  Large-a Altmetrics focuses on social media references to research publications, such as tweets and blog posts.  The analysis explored here uses Altmetrics' tweets data to examine the social media impact of CSS publications.  

```{r altmetrics-data, child = 'altmetrics-data.Rmd', eval = TRUE}
## Data ##
```

## Methods ##

Given the set of DOIs for the target papers, a Python script queries the Altmetrics API, retrieving data on every tweet that references one of the target papers.  The analysis then considers the number of tweets per paper, the number of CSS-related tweets per Twitter account, the estimated reach per paper and over time, and the delay (time between publication and first tweet) and lifespan (time between first and last tweet) for each Tweeted paper.  

```{r altmetrics-analysis, child = 'altmetrics-analysis.Rmd'}
## Results ##
```

## Discussion ##

Data issues — specifically, acquiring a comprehensive list of all and only CSS publications — presented a major challenge for this analysis.  If EPA chooses to explore the use of altmetrics in the future, it may be worthwhile to expand the capabilities of STICS (or a similar system) to track publication DOIs.  For example, after a publication has completed clearance, STICS could send quarterly or biannual reminders to authors, encouraging them to provide final publication metadata (at a minimum, accepted journal and DOI) for their research products.  

Only about half of all CSS papers (`r nrow(papers)/nrow(dataf) * 100`%) received any tweets at all, and many of these papers received only a single tweet.  The distribution of tweets, delay, and lifespan was highly skewed with a long right tail; that is, while many papers receive a few tweets for a short period around the time they are published, a few papers have an especially high number of tweets, a long delay before receiving their first tweet, or are tweeted about for a long period of time.  Only a single paper received more than 15 tweets.  At the same time, these tweets have a large reach, on the order of 100-10,000 people per paper.  

J. Britt Holbrook (personal communication) has proposed that researchers could use altmetrics to identify future opportunities to increase the social impact of their work.  Tables \ref{tab.high_tweet_papers} and \ref{tab.high_tweeters} may be especially useful here.  Table \ref{tab.high_tweet_papers} may indicate which topics or CSS project areas tend to attract relatively broad attention on social media; strategically tweeting about these topics or project areas in the future could be an effective way to increase the social media presence of CSS research.  Similarly, table \ref{tab.high_tweeters} indicates Twitter accounts/users who tend to pay attention to CSS research.  CSS communications staff could engage with these accounts/users, encouraging them to tweet about CSS research more in the future.  

