---
output:
  html_document: default
  pdf_document:
    fig_caption: yes
---
```{r eval = TRUE, echo=FALSE, message=FALSE, warning=FALSE}
fig.height = 2.5
fig.width = 1.5 * fig.height
font_size = 8

knitr::opts_chunk$set(fig.height = fig.height, fig.width = fig.width, 
					  echo = FALSE, message=FALSE, warning=FALSE)

```

# Study 1: Altmetrics #

The term "altmetrics" has two distinct meanings.  Small-a "altmetrics" is a portmanteau of "alternative bibliometrics," that is, alternatives to such metrics as paper citation counts, journal impact factors, and researcher H-indices.  These standard bibliometrics are designed to measure the impact of research within the scholarly community, and thus are generally not useful for assessing social impacts.  Consequently, there is a substantial amount of active work on developing altmetrics that are relevant to social impact.  *[cite Britt etc.]*

Large-a "Altmetrics" refers to "Altmetric.com," a UK-based company that develops an integrated set of small-a altmetrics and makes them available programmatically using a web-based API [application programming interface].  Large-a Altmetrics focuses on social media references to research publications, such as tweets and blog posts.  The analysis explored here uses Altmetrics' tweets data to examine the social media impact of CSS publications.  

```{r altmetrics-data, child = 'altmetrics-data.Rmd', eval = TRUE}
## Data ##
```

## Methods ##

Given the set of DOIs for the target papers, a Python script queries the Altmetrics API, retrieving data on every tweet that references one of the target papers.  The analysis then considers the number of tweets per paper, the number of CSS-related tweets per Twitter account, the estimated reach per paper and over time, and the delay (time between publication and first tweet) and lifespan (time between first and last tweet) for each Tweeted paper.  

*[reference set]*

```{r altmetrics-analysis, child = 'altmetrics-analysis.Rmd'}
## Results ##
```