---
title: 'CompImpact: Computational Tools for Assessing the Social Impact of CSS Research'
author: "Daniel J. Hicks"
output: 
    pdf_document:
        fig_caption: true
        number_sections: true
        toc: true
geometry: margin=1.5in
tables: yes
bibliography: /Users/dhicks1/Dropbox/latex/texmf/bibtex/bib/library.bib
---

```{r general_setup, echo = FALSE, warning = FALSE, messages = FALSE}
fig.height = 2
fig.width = 1.5 * fig.height
font_size = 8

knitr::opts_chunk$set(dev = 'tikz',
			   fig.height = fig.height, fig.width = fig.width,
			   fig.align='center', fig.pos = 't',
			   sanitize = TRUE,
			   echo = FALSE,
			   warning = FALSE, message = FALSE,
			   cache = TRUE
			   )
options(digits = 2, xtable.comment = FALSE, xtable.caption.placement = 'top', 
		xtable.include.rownames = FALSE, 
		xtable.table.placement = 't')

## After http://stackoverflow.com/questions/18965637/set-global-thousand-separator-on-knitr
knitr::knit_hooks$set(inline = function(x) {
	if (is.numeric(x)) {
		## If we want to use scientific notation, pass it to knitr's formatting function
		if ((abs(x) > 10^6)|abs(x) < 10^-4) {
			knitr:::format_sci(x, 'latex')
		} else {
			## Otherwise use format to get thousands separators
			format(x, big.mark = ',', scientific = FALSE, digits = 2)
		}
	} else {
		x
	}
})
```


# Executive Summary #

*[write this]*

- printed/electronic speech acts are easy to study, because by definition they leave tangible records that can be gathered with relatively little effort
- effects on the audience are more difficult to study, because 
	- there are not necessarily tangible records of their reactions, etc.
	- the audience itself may not be well-defined
	- the goals of communication may not be well-defined
	- who are you trying to influence, and what do you want them to do differently? 
	
	
- to work within the communication status quo described in this report:
	- encourage high-tweeting accounts to tweet more
	- strategically emphasize topics that receive more attention, such as the way CSS research will be used under the new TSCA
- to push against the communication status quo:  
	- encourage a more diverse range of tweeters to tweet about CSS research
	- strategically emphasize topics that are underdisseminated, such as the smaller roles CSS research is playing through EPA
	
*[what happened to that l'oreal story?]*


# Introduction #

Over the past forty years, *[something about a shift to research that has a social impact; maybe ref Silver Book?]*.  While researchers at the EPA have always been dedicated to research that aims to protect human health and the environment, they have placed less emphasis on developing empirical methods to measure and assess the social impacts of their work.  

For many areas of EPA research, there are established pathways for deploying research products in regulatory contexts.  The social impact of these research areas can be measured simply in terms of the impacts of the downstream regulation. Conventional air quality research or chemical risk assessments might be good examples of this "normal regulatory science."  However, other research areas — such as research that is actively developing new tools for public health or high-throughput toxicology — do not yet have established regulatory connections or simple regulatory metrics.  (The social impact of a screening prioritization decision for a potential endocrine disruptor, for example, is itself difficult to define.)  Assessing the social impact of these research areas requires more subtle tools.  

This report explores two quantitative approaches to assessing the social impact of CSS research.  Both approaches rely on data-intensive, computational tools.  The report examines data and model-based challenges to deploying and interpreting the outputs of these tools; besides reporting some preliminary findings, this report also offers suggestions for how EPA can improve its ability to use these tools to assess and improve the social impact of its research.  

*[more details]*

*[note on reproducibility]*

```{r, child = 'altmetrics.Rmd'}
```

```{r, child = 'BNA.Rmd'}
```

# References #
